Analysis of the inference of ratings and rankings on Higher Order Networks with complex topologies

### 2025-09-06 draft/draft.tex
progress

A natural approach to inferring rating scores consistent with the pairwise comparisons is to assign ratings $w_i$ (for $i=0,1,\dots,n$) that minimize the sum of squared discrepancies
$$
\sum_{i<j} \bigl(f_{ij}-(w_i-w_j)\bigr)^2.
$$
This yields a well-posed optimization problem that admits an elegant reformulation in the language of cochains over simplicial complexes.
The resulting framework for the inference of ratings and rankings is known as {\em HodgeRank}~\cite{jiang2011statistical, perotti2025analysis}.

@@

>>>>

A natural approach to inferring rating scores consistent with the pairwise comparisons is to assign ratings $f_i$ (for $i=0,1,\dots,n$) that minimize the sum of squared discrepancies
$$
\sum_{i<j} \bigl(f_{ij}-(f_i-f_j)\bigr)^2.
$$
This yields a well-posed optimization problem that admits an elegant reformulation in the language of cochains over simplicial complexes.
The resulting framework for the inference of ratings and rankings is known as {\em HodgeRank}~\cite{jiang2011statistical, perotti2025analysis}.

@@

Any $k$-cochain $f$ admits the so called Hodge decomposition
$$
f = g\oplus h\oplus s
$$
where the direct sum components $g\in \rng d_{k-1}$, $h\in \krn L_k$ and $s\in \rng d_k^*$ are unique and are called the gradient, the harmonic and the solenoid components, respectively.
As a consequence, there exists non-unique cochains $w\in C_{k-1}^*$ and $u\in C_{k+1}^*$ such that $g=d_{k-1}w$ and $s=d_k^*u$, that can be found solving for
\begin{eqnarray}
d_kf 
&=&
d_k(g+h+s)
\notag
\\
&=&
d_k s
\notag
\\
&=&
d_k d_k^* u
\notag
\\
&=&
L_{k+1}^{\downarrow} u
\notag
\end{eqnarray}
and
\begin{eqnarray}
d_{k-1}^* f 
&=&
d_{k-1}^*(g+h+s)
\notag
\\
&=&
d_{k-1}^* g
\notag
\\
&=&
d_{k-1}^*d_{k-1} w
\notag
\\
&=&
L_{k-1}^{\uparrow} w,
\notag
\end{eqnarray}
while the harmonic component can be found from $h=f-g-s$.

\subsection{Discrete calculus reformulation of the infererence of ratings}
